---
title: "Analysis and estimation of factors associated with Obesity"
subtitle: 'STAT 420, Final Project Proposal'
author: 
- "Adam Goodell"
- "Joel Dazo"
- "Rahul Sharma"
date: ''
geometry: margin=2cm
output:
  html_document: 
    theme: readable
    toc: yes
  pdf_document: default

urlcolor: cyan
link-citations: yes
---
```{css, echo=FALSE}
h1,h3 {
  text-align: center;
}
```

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
options(scipen = 1, digits = 4, width = 80, fig.align = "center")
knitr::opts_chunk$set(echo = TRUE)
library(RWeka)
```

***

## Data file description

We will be using obesity dataset to analyze and estimate the factors associated with the obesity. 

The aim of this dataste is to provide an estimate of the obesity levels of persons from the countries of Mexico, Peru and Colombia [1]. It contains the information about two principal areas were captured through an online survey in order to accomplish this objective:  the dietary habits and the behavioral characteristics of individuals.  Attributes about dietary habits include: frequent consumption of high caloric food (FAVC), frequency of consumption of vegetables (FCVC), among others. Attributes about behavioral characteristics include: calorie consumption monitoring (SCC), physical activity frequency (FAF), time using technology devices (TUE), among others. It also contains physical characteristics such as gender, age, height and weight.  The dataset had a combination of numeric and mostly nominal categorical attributes.

The obesity data is represented as comma-separated values with `17` variables or attributes and `2,111` records. It's made available in CSV format as well as in an Attribute-Relation File Format (ARFF) format for use with the Weka machine learning software. *Our primary variable of interest for predictions is Obseity levels and weight, given some of the attributes stated above.*

The dataset can be accessed and downloaded by following this <a href="https://urldefense.com/v3/__https://archive.ics.uci.edu/ml/machine-learning-databases/00544/ObesityDataSet_raw_and_data_sinthetic*20(2).zip__;JQ!!DZ3fjg!smWxBizt9_P1O4hbNghVvRKkrJtci8J3d-wBQp8u_g8sio-T1OfpK5ekJP97MVyXt4fO$">link</a>.

***

## Dataset background details

The initial dataset was collected by author [1] from online survey respondents who answered dietary, physical and behavioral questions. The data underwent data cleaning (removed missing or incomplete data, dealt with atypical data, data normalization, etc.). It also went through a data engineering process where an interim Body Mass Index value was calculated and compared with external World Health Organization (WHO) data to augment the dataset with the obesity level categorical attribute.

The labeling process unearthed and identified a problem concerning unbalanced categories (balancing class problem) to which the use of the Weka tool and the application of Synthetic Minority Over-Sampling Technique (<a href="https://urldefense.com/v3/__https://en.wikipedia.org/wiki/Oversampling_and_undersampling_in_data_analysis*SMOTE__;Iw!!DZ3fjg!smWxBizt9_P1O4hbNghVvRKkrJtci8J3d-wBQp8u_g8sio-T1OfpK5ekJP97MRwWo5h-$">SMOTE</a>) generated an additional 1,626 synthetic records to resolve the balancing class issue. Including the synthetic records, there are 2,111 records in the dataset.

```{r}
obesity=read.arff("Obesity.arff"); dim(obesity)
```

```{r}
head(obesity[, c(1,2,3,4,6,7,8,9,10,11,12)], 7)
```

```{r Show Observations}
obesity$Weight[1:10]
```

***

## Motivations

The goal of this project is to analyze and estimate the factors associated with the obesity. Obesity is one of the major problems in the current generation and has been growing steadily. There are many scientific publications over the past few years, specifically concentrating on the factors included in this project's dataset that is used largely in software industry to build better tools. 

We chose this research topic after analyzing the facts from WHO about obesity [2]:

- Worldwide obesity has nearly tripled since 1975.
- In 2016, more than 1.9 billion adults, 18 years and older, were overweight. Of these over 650 million were obese.
- 39% of adults aged 18 years and over were overweight in 2016, and 13% were obese.
- Most of the world's population live in countries where overweight and obesity kills more people than underweight.
- 39 million children under the age of 5 were overweight or obese in 2020.
- Over 340 million children and adolescents aged 5-19 were overweight or obese in 2016.

The source our research project is the paper published by Fabio Mendoza Palechor, this paper presents data for the estimation of obesity levels in individuals from the countries of Mexico, Peru and Colombia [1] and we acquired the dataset from the UCI Machine Learning Repository [3]. The dataset contains numerical, ordinal and categorical attributes and has sufficient observations for regression study that fulfills this course requirement.

***Primarily our aim is to explore these topics but not limited to***

- Estimation of the obesity level of individuals using various eating habits.
- Estimation of the obesity level of individuals using various physical conditions.
- How other variables are associated with the Obesity Gender, Age, Height and Weight impacts.
- Build the model to predict Weights and Obesity levels using linear regression and logistic regression respectively.

In addition to predicting weight, our research metrics about obesity and body mass index (BMI) shall be compared with the baseline published by Mendoza and WHO [1]:

- Underweight Less than 18.5 
- Normal 18.5 to 24.9
- Overweight 25.0 to 29.9
- Obesity I 30.0 to 34.9
- Obesity II 35.0 to 39.9
- Obesity III Higher than 40

***

## References
[1]. Fabio Mendoza Palechor; Alexis de la Hoz Manotas, Dataset for estimation of obesity levels based on eating habits and physical condition in individuals from Colombia, Peru and Mexico. Data in Brief, Volume 25, 2019,104344.

[2]. <a href="https://www.who.int/news-room/fact-sheets/detail/obesity-and-overweight">Obesity and overweight</a>

[3]. <a href="https://archive.ics.uci.edu/ml/datasets/Estimation+of+obesity+levels+based+on+eating+habits+and+physical+condition+">Estimation of obesity levels based on eating habits and physical condition Data Set, UCI Machine Learning Repository</a>

[4]. Eduardo De-La-Hoz-Correa, Fabio E. Mendoza-Palechor, Alexis De-La-Hoz-Manotas, Roberto C. Morales-Ortega and Sánchez Hernández Beatriz Adriana, Obesity Level Estimation Software based on Decision Trees

[5]. Obesity: a public health problem, MAGAZINE OF SCIENTIFIC AND TECHNOLOGICAL DISCLOSURE OF THE VERACRUZANA UNIVERSITY.

[6]. <a href="https://www.nhlbi.nih.gov/health/educational/lose_wt/BMI/bmicalc.htm">Body Mass Index calculator</a>

